---
title: "Get Started with covdata"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{covdata}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Loading the Package

The `covdata` package aims to make data related to the COVID-19 pandemic easily accessible to users of R. Once the package is installed, load it in the usual way:

```{r setup}
library(tidyverse) 
library(covdata)
```

Loading the package makes a variety of datasets available for use. Because the data are in tibbles, the use of the `tidyverse` suite of packages is strongly recommended, though it is not required. If use the data objects as dataframes (i.e., without loading the `tidyverse` packages) you will need to re-encode some variables, most importantly `date` and `date-time` columns, for them to behave as expected.

## Exploring the Data

Individual datasets are documented on their help pages. The documentation includes details on properly citing the source of the data and, in some cases, the terms of use under which it is provided. The vignettes in the package provide a brief overview of several of the datasets. These include:

- For the ECDC data: `vignette("ecdc-data")`
- For the US data: `vignette("us-data")`
- For the New York Times data: `vignette("new-york-times")`
- For the global mobility data: `vignette("mobility-data")`
- For the CoronaNet Project data: `vignette("coronanet-data")`

## Other Data Sources

- [Philip Cohen](https://familyinequality.wordpress.com) has released several datasets and code for STATA users. These can be found [at his website](https://familyinequality.wordpress.com/2020/04/27/covid-19-code-data-codebooks-figures/).
